---
output:
  html_document: default
  pdf_document: default
---
## Title: "Modeling Coral Invertebrate Communities Using 3D photogrammetry"
author: "Journ Galvan"
date: "4/23/2020"
output: html_document
---

```{r, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
# Load libraries
library(here)
library(tidyverse)
library(vegan)
library(reshape)
library(ggplot2)
library(ggpubr)
library(zoo)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(psych)
library(knitr)
library(faraway)
library(car)
library(MASS)
library(gridExtra)
library(grid)
```

```{r}
#Load data 
branch_width <- read.csv(here("photogrammetry_data/branchwidth_data.csv"))
coral_pg <- read.csv(here("photogrammetry_data/galvan_journ_datasheet_v2.csv"))
coral_field <- read.csv(here("photogrammetry_data/field_experiment_colony_measurements_moorea_summer2019.csv"))
cafi_surveys <- read.csv(here("cafi_data/revised_cafi_data_moorea_summer2019_11_27.csv"))
cafi_field <- read.csv(here("cafi_data/prelim_cafi_counts_moorea_summer2019.csv"))
updated_cafi <- read.csv(here("cafi_data/cafi_data_w_taxonomy_summer2019_2020_5_21.csv"))
```


```{r}
#Incorporate cleaned CAFI data
updated_cafi2 <- updated_cafi %>% filter(str_detect(coral_id, "^FE")) %>%
  dplyr::select(master_sort, coral_id, code, type, search_term, lowest_level, phylum, genus, species, general_notes) %>%
  subset(phylum!="Chordata")
```

```{r, warning=FALSE}
#Calculate CAFI richness, abundance, and diversity (shannon weiner) for each coral 
cafi_summarized2 <- group_by(updated_cafi2, coral_id) %>%
  summarise(num_cafi = n(), cafi_richness = length(unique(code)), cafi_present = paste(sort(unique(code)), collapse = ";"))
cafi_summarized2$sw <- updated_cafi2 %>% 
  count(code, coral_id = coral_id) %>% 
  spread(code,n) %>% 
  mutate_all(list(~tidyr::replace_na(.,0))) %>% 
  dplyr::select(-coral_id) %>% 
  diversity(index = "shannon")
```

```{r}
### Incorporate coral data 

# Clean field data
coral_field2 <- coral_field %>%
  dplyr::rename(branch = branch_width)

# Clean photogrammetry morphometric data
coral_pg$volume_pg <- as.numeric(as.character(coral_pg$volume_pg)) # Convert factor to numeric

# Clean photogrammetry branch width data and take averages of branch distance for each coral
branch_width$branch_distance_mm <- as.numeric(as.character(branch_width$branch_distance_mm)) # Convert factor to numeric
branch_w_summarized <- group_by(branch_width, coral_id) %>%
  summarise(avg_w_cm = sum(branch_distance_mm/10)/n(), #take average branch distance and convert mm to cm
            measurements = length(unique(replicate_measurement)), 
            locations = paste(sort(unique(location)), collapse = ";"))
```

```{r}
#Create dataframe called coral_dim 
coral_dim <- merge(coral_pg, branch_w_summarized, by = "coral_id") %>%
 merge(coral_field2, by = "coral_id") %>%
  mutate(volume_pg=volume_pg*10^6,
         max_hull_volume=max_hull_volume*10^6, #convert m^3 to cm^3
         max_hull_surface_area=max_hull_surface_area*10^4,
         surface_area=surface_area*10^4) %>%   #convert m^2 to cm^2
  dplyr::rename(max_hull_SA = max_hull_surface_area,
                SA = surface_area)
          
coral_dim$interstitial_space <- coral_dim$max_hull_volume - coral_dim$volume_pg #calculate available space by subtracting software estimated volume from convex hull volume

coral_dim$SAV <- coral_dim$SA / coral_dim$volume_pg #calculate surface area to volume relationship

coral_dim$convexity <- coral_dim$volume_pg / coral_dim$max_hull_volume #calculate proportion occupied, high ratios indicate less free space between branches

coral_dim$packing <- coral_dim$SA / coral_dim$max_hull_SA #how much of an objects surface area is situated internally

coral_dim$sphericity <- ((3.14^(1/3))*((6*coral_dim$volume_pg)^(2/3)))/coral_dim$SA #calculate sphericity or how close the object is to a sphere

coral_dim <- coral_dim %>% 
  dplyr::select(coral_id, branch, cafi, size_class, volume_field, volume_pg, SA, avg_w_cm, interstitial_space, SAV, convexity, packing, sphericity) %>%
  filter(cafi=="empty")

coral_dim$branch <- ifelse(coral_dim$convexity>=0.5, "tight","wide") #classifies wide and tight branching coral based on convexity measurement
```

```{r}
#Create dataframe called cafi_coral and merge coral and cafi data
cafi_coral <- merge(cafi_summarized2, coral_dim, by ="coral_id") %>%
  drop_na(interstitial_space) %>%
  dplyr::select(coral_id, branch, cafi, size_class, volume_field, volume_pg, SA, avg_w_cm, interstitial_space, SAV, convexity, packing, sphericity, num_cafi, cafi_present, cafi_richness, sw) 
```

```{r}
#Log transform variables
cafi_coral$log_num_cafi <- log(cafi_coral$num_cafi)
cafi_coral$log_volume_field <- log(cafi_coral$volume_field)
cafi_coral$log_volume_pg <- log(cafi_coral$volume_pg)
cafi_coral$log_avg_w <- log(cafi_coral$avg_w_cm)
cafi_coral$log_interstitial_space <- log(cafi_coral$interstitial_space)
cafi_coral$log_SA <- log(cafi_coral$SA)
cafi_coral$log_cafi_richness <- log(cafi_coral$cafi_richness)
cafi_coral$log_sphericity <- log(cafi_coral$sphericity)

#Sqrt transform variables
cafi_coral$sqrt_num_cafi <- sqrt(cafi_coral$num_cafi)
cafi_coral$sqrt_volume_pg <- sqrt(cafi_coral$volume_pg)
cafi_coral$sqrt_interstitial_space <- sqrt(cafi_coral$interstitial_space)
cafi_coral$sqrt_SA <- sqrt(cafi_coral$SA)
cafi_coral$sqrt_cafi_richness <- sqrt(cafi_coral$cafi_richness)
cafi_coral$sqrt_sphericity <- sqrt(cafi_coral$sphericity)
```

```{r}
#Visualize correlations between variables
cor_var <- cafi_coral %>% 
  dplyr::select(num_cafi,volume_pg,volume_field,SA,avg_w_cm,interstitial_space,sphericity,convexity,packing) 
pairs.panels(cor_var, 
             method="pearson", 
             hist.col = "white",
             density=TRUE,
             ellipses=FALSE)

#Use log transformed data
cor_var2 <- cafi_coral %>%
  dplyr::select(log_num_cafi,log_volume_pg,log_volume_field,log_SA,log_avg_w,log_interstitial_space,log_sphericity,convexity,packing)
pairs.panels(cor_var2, 
             method="pearson", 
             hist.col = "white",
             density=TRUE,
             ellipses=FALSE)
```

```{r}
g1 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=log_SA, shape=branch, color=branch, linetype=branch))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        legend.position="none")+
  labs(y=expression(paste("Log SA cm"^{2})))

g2 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=log_avg_w, shape=branch, color=branch, linetype=branch))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        legend.position="none")+
  labs(y="Log branch distance cm")

g3 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=log_interstitial_space, shape=branch, color=branch, linetype=branch))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        legend.position="none")+
  labs(y=expression(paste("Log space cm"^{3})))

g4 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=log_sphericity, shape=branch, color=branch, linetype=branch))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        legend.position="none")+
  labs(y="Log sphericity")

g5 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=convexity, shape=branch, color=branch, linetype=branch))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        legend.position="none")+
  labs(y="Convexity", x=expression(paste("Log Vol. cm"^{3})))

g6 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=packing, shape=branch, color=branch, linetype=branch))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(axis.title.x=element_blank(),
        legend.position="bottom")+
  labs(y="Packing")

#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend<-g_legend(g6)
       
fig1 <- ggarrange(g1,g2,g3,g4,g5,g6,
                     ncol = 3, nrow=2)
  
annotate_figure(fig1,
                top = text_grob("Morphological Measurements", color = "black", face = "bold", size = 14),
                bottom = text_grob(expression(paste("Log Vol. cm"^{3})), color = "black",
                                   face = "bold", size = 12),
                fig.lab = "Figure 1", fig.lab.face = "bold")  
```

```{r}
pg_field <- ggplot(cafi_coral, aes(x=volume_pg, y=volume_field))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  scale_y_continuous(breaks=seq(0,max(cafi_coral$volume_field), by = 5000))+
  scale_x_continuous(breaks=seq(0,max(cafi_coral$volume_pg), by = 2000))+
  theme_bw()+
  labs(y= expression(paste("elipsoid V cm"^{3})), x= expression(paste("software V cm"^{3})))
  
pg_field
```

```{r}
invert_field <- ggplot(cafi_coral, aes(x=volume_field, y=num_cafi))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  theme_bw()+
  labs(y="invert abundance", x= expression(paste("elipsoid V cm"^{3})))
  
invert_field

lm_invert_field <- lm(num_cafi~volume_field, data=cafi_coral)

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  volume_field = "elipsoid volume")

tab_model(lm_invert_field,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")
```

```{r}
invert_pg <- ggplot(cafi_coral, aes(x=volume_pg, y=num_cafi))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  theme_bw()+
  labs(y="invert abundance", x= expression(paste("software V cm"^{3})))
  
invert_pg

lm_invert_pg <- lm(num_cafi~volume_pg, data=cafi_coral)

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  volume_pg = "software est. volume")

tab_model(lm_invert_pg,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")
```


```{r, warning=FALSE}
# Boxplot of invert abundance
log_invert_box <- ggplot(cafi_coral, aes(x= branch, y=log_num_cafi, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter()+ # adds data on top of box plot
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(cafi_coral$log_num_cafi + 1)))+
  theme_classic()+
  labs(y="Log invert. abundance", x="Coral branching", title = "Log invert. abundance boxplot")

log_invert_box

# Boxplot of invertebrate species richness per coral for both tight and wide branching
richness_box <- ggplot(cafi_coral, aes(x= branch, y=log_cafi_richness, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter()+
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(cafi_coral$log_cafi_richness + 1)))+
  theme_classic()+
  labs(y="Invert. sp. richness", x="Coral branching", title = "Invert. sp. richness boxplot")

richness_box

# Boxplot of invertebrate diversity for both tight and wide branching coral
diversity_box <- ggplot(cafi_coral, aes(x= branch, y=sw, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter()+
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(cafi_coral$sw + 0.5)), breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3))+
  theme_classic()+
  labs(y="Invert. diversity", x="Coral branching", title = "Invert. diversity boxplot")

diversity_box
```

```{r}
#Q-Q plot separated by group 
with(cafi_coral, qqPlot(num_cafi[branch == "wide"], dist="norm")) 
with(cafi_coral, qqPlot(num_cafi[branch == "tight"], dist="norm")) 

#Shapiro Wilk seperated by group
with(cafi_coral, shapiro.test(num_cafi[branch == "wide"])) 
with(cafi_coral, shapiro.test(num_cafi[branch == "tight"])) 
```
With a $p>0.05$ we fail to reject the null hypothesis that our data is normally distributed for wide branching corals. With a $p<0.05$ we reject the null hypothesis that our data is normally distributed for tight branching corals. Because we are working with smaller data sets, we cannot use the central limit theorem. 

```{r}
#Levene's test: response variable, group variable
leveneTest(cafi_coral$num_cafi, cafi_coral$branch)
```
With a $p>0.05$ we fail to reject the null hypothesis that our data is normally distributed for tight and wide branching corals. Both groups of corals that have tight and wide branching morphologies have equal variance.

Because tight branching coral is not normally distributed, we could either perform a transformation or non-parametric test. We will try log transforming our data first. A column log transforming the number of inverts has already been created

```{r}
#Q-Q plot separated by group 
with(cafi_coral, qqPlot(log_num_cafi[branch == "wide"], dist="norm")) 
with(cafi_coral, qqPlot(log_num_cafi[branch == "tight"], dist="norm")) 

#Shapiro Wilk seperated by group
with(cafi_coral, shapiro.test(log_num_cafi[branch == "wide"])) 
with(cafi_coral, shapiro.test(log_num_cafi[branch == "tight"])) 
```
```{r}
#Levene's test: response variable, group variable
leveneTest(cafi_coral$log_num_cafi, cafi_coral$branch)
```

```{r}
# subset data
tight <- subset(cafi_coral, branch == "tight")
wide <- subset(cafi_coral, branch == "wide")

t.test(tight$log_num_cafi, wide$log_num_cafi, var.equal = TRUE)
```
With a $p>0.05$ we fail to reject the null hypothesis that the difference between the means of the log transformed invert counts are different from zero. 

##### Since there does not appear to be a signficant difference in invertebrate counts between tight and wide corals, we will now test how well our software estimated method of calculating volume and interstitial space relates to invertebrate counts.

We will first assess $p<0.05$ to test if the linear regression slope was significantly different from zero. Correlation of dependent and independent variables $R^2$ was also included in the tables. 

```{r, message=FALSE}
#Create graphs of log transformed invertebrate abundance
log_invert_pg <- ggplot(cafi_coral, aes(x=log_volume_pg, y=log_num_cafi))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE)+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="log invert. abundance", x= expression(paste("log software est. V cm"^{3})))

log_invert_pg

#Residuals for log transformed data
lm_log_invert_pg <- lm(log_num_cafi~log_volume_pg, data=cafi_coral)
par(mfrow=c(2,2))
plot(lm_log_invert_pg)
par(mfrow=c(1,1))

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  log_volume_pg = "Log software est. V"
)
tab_model(lm_log_invert_pg,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")
```

```{r}
log_invert_field <- ggplot(cafi_coral, aes(x=log_volume_field, y=log_num_cafi))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE)+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="log invert. abundance", x= expression(paste("log elipsoid V cm"^{3})))

log_invert_field

lm_log_invert_field <- lm(log_num_cafi~log_volume_field, data=cafi_coral)
par(mfrow=c(2,2))
plot(lm_log_invert_field)
par(mfrow=c(1,1))

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  log_volume_field = "Log elispoid V"
)
tab_model(lm_log_invert_field,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")
```


```{r, message=FALSE}
#Create graphs of log transformed invertebrate abundance
log_cafi_space <- ggplot(cafi_coral, aes(x=log_interstitial_space, y=log_num_cafi))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE)+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Log invert. abundance", x= expression(paste("Log interstitial space cm"^{3})), title = "Log 3D available space and inverts")

log_space_rich <- ggplot(cafi_coral, aes(x=log_interstitial_space, y=log_cafi_richness))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE)+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Log sp. richness", x= expression(paste("Log interstitial space cm"^{3})), title = "Log available space and sp. richness")

fig2 <- ggarrange(log_cafi_space, log_space_rich,
                     ncol = 2, nrow = 1) #combines all graphs
fig2

#Residuals for log transformed data
log_lm_space <- lm(log_num_cafi~log_interstitial_space, data=cafi_coral)
par(mfrow=c(2,2))
plot(log_lm_space)
par(mfrow=c(1,1))

log_lm_rich <- lm(log_cafi_richness~log_interstitial_space, data=cafi_coral)
par(mfrow=c(2,2))
plot(log_lm_rich)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  log_interstitial_space = "Log interstitial space"
)
tab_model(log_lm_space, log_lm_rich,
          pred.labels = pl,
          dv.labels = c("Log invertebrate abundance", "Log species richness"))
```

Invertebrate abundance and species richness showed $p<0.05$ for both volume and interstitial space. We reject the null hypothesis that the slope of the regression line is equal to zero. The correlations for all four linear regressions were $R^2>30$

##### We will now fit a single variable Yi=β0+β1Xi+ϵi or multivariable linear model Y=β0+β1X1+β2X2+...+βnXn+ϵ to our data using different combinations of explanatory variables. Our response variable will be abundance of invertebrates. We will use AIC and R-squared values to determine the best fit model. 

```{r}
#Visualize correlations between variables
# subset data
log_tight <- cafi_coral %>%
  subset(branch == "tight") %>%
  dplyr::select(log_num_cafi,log_volume_pg,log_SA,log_interstitial_space,log_sphericity,convexity,packing)
pairs.panels(log_tight, 
             method="pearson", 
             hist.col = "white",
             density=TRUE,
             ellipses=FALSE)

log_wide <- cafi_coral %>%
  subset(branch == "wide") %>%
  dplyr::select(log_num_cafi,log_volume_pg,log_SA,log_interstitial_space,log_sphericity,convexity,packing)
pairs.panels(log_wide, 
             method="pearson", 
             hist.col = "white",
             density=TRUE,
             ellipses=FALSE)

log_cafi_coral <- cafi_coral %>%
  dplyr::select(log_num_cafi,log_volume_pg,log_SA,log_interstitial_space,log_sphericity,convexity,packing)
pairs.panels(log_cafi_coral, 
             method="pearson", 
             hist.col = "white",
             density=TRUE,
             ellipses=FALSE)
```

```{r}
plot1 <- ggplot(cafi_coral, aes(x=log_volume_pg, y=log_num_cafi))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  theme_classic()+
  theme(axis.title.y=element_blank(),
        legend.position="none")+
  labs(x=expression(paste("Log software est. V cm"^{3})))


plot2 <- ggplot(cafi_coral, aes(x=log_volume_field, y=log_num_cafi))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  theme_classic()+
  theme(axis.title.y=element_blank(),
        legend.position="none")+
  labs(x=expression(paste("Log elipsoid V cm"^{3})))


plot3 <- ggplot(cafi_coral, aes(x=log_sphericity, y=log_num_cafi))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  theme_classic()+
  theme(axis.title.y=element_blank(),
        legend.position="none")+
  labs(x="Log sphericity")

plot4 <- ggplot(cafi_coral, aes(x=packing, y=log_num_cafi))+ 
  geom_point()+
  geom_smooth(method=lm, se=TRUE)+
  theme_classic()+
  theme(axis.title.y=element_blank(),
        legend.position="none")+
  labs(x="Packing")

fig3 <- ggarrange(plot1,plot2,plot3,plot4,
                     ncol = 2, nrow=2)
  
annotate_figure(fig3,
                top = text_grob("Morphological Measurements", color = "black", face = "bold", size = 14),
                left = text_grob("Log invert. abundance", color = "black",
                                   rot = 90, face = "bold", size = 12),
                fig.lab = "Figure 1", fig.lab.face = "bold") 
```

```{r}
#Linear model for log sphericity and invert abundance
lm_sphericity <- lm(log_num_cafi~log_sphericity, data=cafi_coral)
par(mfrow=c(2,2))
plot(lm_sphericity)
par(mfrow=c(1,1))

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  log_sphericity = "Log sphericity"
)
tab_model(lm_sphericity,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")

#Linear model for packing and invert abundance
lm_packing <- lm(log_num_cafi~packing, data=cafi_coral)
par(mfrow=c(2,2))
plot(lm_packing)
par(mfrow=c(1,1))

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
   packing = "Packing"
)
tab_model(lm_packing,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")

#Linear model for convexity and invert abundance
lm_convexity <- lm(log_num_cafi~convexity, data=cafi_coral)
par(mfrow=c(2,2))
plot(lm_convexity)
par(mfrow=c(1,1))

#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
   convexity = "Convexity"
)
tab_model(lm_convexity,
          pred.labels = pl,
          dv.labels = "Invertebrate abundance")
```

Log transformation appears normal. Log transformed volume, surface area, interstitial space, and packing were morphological measurements that had a positive correlation to number of cafi. Log transformed sphericity has a negative correlation to cafi. Because our data does not contain any 0 counts for invertebrate (meaning every coral contained an invertebrate population) we applied a log transformation to our count data. We will compare our log transformed poisson general linear model with a quasi-poisson general linear model with non-log transformed counts. 

VIF>10 will be removed from the dataset 

```{r}
lm1 <- lm(log_num_cafi~log_volume_pg+log_sphericity+packing, data=log_cafi_coral)
vif(lm1) 

lm_tight <- lm(log_num_cafi~log_volume_pg+packing, data=log_tight)
vif(lm_tight)

lm_wide <- lm(log_num_cafi~log_volume_pg+log_sphericity+packing, data=log_wide)
vif(lm_wide)
```

```{r}
res1=lm1$residuals

#Test normality assumption on residuals
par(mfrow=c(2,2))
plot(lm1)
par(mfrow=c(1,1))

hist(res1)
shapiro.test(res1)

#Data is FINALY! normal
```

```{r}
fullmodel <- lm(log_num_cafi~log_volume_pg+log_sphericity+packing, data=log_cafi_coral)
nullmodel <- lm(log_num_cafi~1, data=log_cafi_coral)

stepAIC(fullmodel, scope = c(upper = fullmodel, lower = nullmodel), direction = "both")
```

```{r}
fullmodel_tight <- lm(log_num_cafi~log_volume_pg+log_sphericity+packing, data=log_tight)
nullmodel_tight <- lm(log_num_cafi~1, data=log_tight)

stepAIC(fullmodel_tight, scope = c(upper = fullmodel_tight, lower = nullmodel_tight), direction = "both")
```

```{r}
fullmodel_wide <- lm(log_num_cafi~log_volume_pg+log_sphericity+packing, data=log_wide)
nullmodel_wide <- lm(log_num_cafi~1, data=log_wide)

stepAIC(fullmodel_wide, scope = c(upper = fullmodel_wide, lower = nullmodel_wide), direction = "both")
```


