---
output:
  html_document: default
  pdf_document: default
---
## Title: "Updated Linear regression of corals and inverts"
author: "Journ Galvan"
date: "4/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r load libraries, include=FALSE}
#Code referenced from lines 83-125 in "Coral Condition Data Preliminary Analysis - Stier Lab Moorea Summer 2019" Author: Joseph Curtis

# Load libraries
library(here)
library(tidyverse)
library(vegan)
library(reshape)
library(ggplot2)
library(ggpubr)
library(zoo)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(psych)
library(knitr)
```

```{r load data, include=FALSE}
# Load data
branch_width <- read.csv(here("photogrammetry_data/branchwidth_data.csv"))
coral_pg <- read.csv(here("photogrammetry_data/galvan_journ_datasheet_v2.csv"))
coral_field <- read.csv(here("photogrammetry_data/field_experiment_colony_measurements_moorea_summer2019.csv"))
cafi_surveys <- read.csv(here("cafi_data/revised_cafi_data_moorea_summer2019_11_27.csv"))
cafi_field <- read.csv(here("cafi_data/prelim_cafi_counts_moorea_summer2019.csv"))
updated_cafi <- read.csv(here("cafi_data/cafi_data_w_taxonomy_summer2019_2020_5_21.csv"))
```


```{r, include=FALSE}
# Incorporate cleaned CAFI data
updated_cafi2 <- updated_cafi %>% filter(str_detect(coral_id, "^FE")) %>%
  select(master_sort, coral_id, code, type, search_term, lowest_level, phylum, genus, species, general_notes) %>%
  subset(phylum!="Chordata")
```

```{r, include=FALSE}
# Calculate CAFI richness, abundance, and diversity (shannon weiner) for each coral 
cafi_summarized2 <- group_by(updated_cafi2, coral_id) %>%
  summarise(num_cafi = n(), cafi_richness = length(unique(code)), cafi_present = paste(sort(unique(code)), collapse = ";"))
cafi_summarized2$sw <- updated_cafi2 %>% 
  count(code, coral_id = coral_id) %>% 
  spread(code,n) %>% 
  mutate_all(list(~tidyr::replace_na(.,0))) %>% 
  select(-coral_id) %>% 
  diversity(index = "shannon")
```

```{r, include=FALSE}
### Incorporate coral data 

# Clean field data
coral_field2 <- coral_field %>%
  dplyr::rename(branch = branch_width)

# Clean photogrammetry morphometric data
coral_pg$volume_pg <- as.numeric(as.character(coral_pg$volume_pg)) # Convert factor to numeric

# Clean photogrammetry branch width data and take averages of branch distance for each coral
branch_width$branch_distance_mm <- as.numeric(as.character(branch_width$branch_distance_mm)) # Convert factor to numeric
branch_w_summarized <- group_by(branch_width, coral_id) %>%
  summarise(avg_w_cm = sum(branch_distance_mm/10)/n(), #take average branch distance and convert mm to cm
            measurements = length(unique(replicate_measurement)), 
            locations = paste(sort(unique(location)), collapse = ";"))
```

```{r, include=FALSE}
#Compare photogrammetry and manual measurements
coral_dim <- merge(coral_pg, branch_w_summarized, by = "coral_id") %>%
 merge(coral_field2, by = "coral_id") %>%
  mutate(volume_pg=volume_pg*10^6,
         max_hull_volume=max_hull_volume*10^6, #convert m^3 to cm^3
         surface_area=surface_area*10^4) #convert m^2 to cm^2

coral_dim$interstitial_space <- coral_dim$max_hull_volume - coral_dim$volume_pg #calculate available space by subtracting software estimated volume from convex hull volume

coral_dim <- coral_dim %>% select(coral_id, branch, cafi, size_class, volume_field, volume_pg, surface_area, avg_w_cm, interstitial_space, max_hull_volume) %>%
  filter(cafi=="empty")

cafi_coral2 <- merge(cafi_summarized2, coral_dim, by ="coral_id") %>%
  drop_na(interstitial_space)
```


### Correlating morphological measurements

##### For this step we related coral volume to surface area and avgerage branch width measurements. Summary data from the linear regression was extracted and exported to a table using the *tab_model( )* function and was then applied to all other linear models in the R markdown document. 

##### We first assesed $p<0.05$ to test if the linear regression slope was significantly different from zero. Correlation of dependent and independent variables $R^2$ was also included in the tables. If the slopes were signficiant, we compared confidence intervals for overlap to give a conservative estimate if the slopes were signficantly different from eachother. Linear regressions for both tight and wide branching corals were overlayed on the same graph to show differences in their slope and standard error was included around the line for visual interpretation.

```{r, message=FALSE}
# Relating coral volume to average branch distance
vol_branch <- ggplot(cafi_coral2, aes(x=volume_pg, y= avg_w_cm, shape= branch, color= branch, linetype= branch))+ #changed 'group' to 'shape'
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  theme(legend.position = "none")+
  labs(y= "Avg. branch distance mm", x= expression(paste("Volume m"^{3}), title= "Volume vs. avg. branch distance"))

vol_SA <- ggplot(cafi_coral2, aes(x=volume_pg, y= surface_area, shape= branch, color= branch, linetype= branch))+ #changed 'group' to 'shape'
  geom_point()+
  geom_smooth(method = lm, se=TRUE, show.legend = FALSE, aes(fill=branch))+
  theme_classic()+
  theme(legend.position = "right")+
  labs(y= expression(paste("Surface Area m"^{2})), x= expression(paste("Volume m"^{3}), title= "Volume vs. SA"))

#fig1 <- vol_branch + vol_SA
#fig1

fig1 <- ggarrange(vol_branch, vol_SA,
                     ncol = 2, nrow = 1) #combines all graphs
fig1
```

```{r, include=FALSE}
# fit linear model to tight coral volume and branch width
fit_vol_branch_tight <- lm(avg_w_cm~volume_pg, data=cafi_coral2, subset=(branch=="tight"))
```

```{r, include=TRUE}
par(mfrow=c(2,2))
plot(fit_vol_branch_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE}
# fit linear model to wide coral volume and branch width
fit_vol_branch_wide <- lm(avg_w_cm~volume_pg, data=cafi_coral2, subset=(branch=="wide"))
```

```{r, include=TRUE}
par(mfrow=c(2,2))
plot(fit_vol_branch_wide)
par(mfrow=c(1,1))
```

```{r}
# HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  volume_pg = "Volume"
)
tab_model(fit_vol_branch_tight, fit_vol_branch_wide,
          pred.labels = pl,
          dv.labels = c("Tight branching avg. distance", "Wide branching avg. distance"))
```

```{r, include=FALSE}
# fit linear model to tight coral volume and surface area
fit_vol_SA_tight <- lm(avg_w_cm~volume_pg, data=cafi_coral2, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_vol_SA_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE}
# fit linear model to wide coral volume and surface area
fit_vol_SA_wide <- lm(surface_area~volume_pg, data=cafi_coral2, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_vol_SA_wide)
par(mfrow=c(1,1))
```

```{r}
# HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  volume_pg = "Volume"
)
tab_model(fit_vol_SA_tight, fit_vol_SA_wide,
          pred.labels = pl,
          dv.labels = c("SA Tight", "SA Wide"))
```

##### The slope is sig. for volume and avg. branch distance as well as the volume and SA of wide corals but not for tight branching. Wide branching also shows a stronger correlation for both morphological comparisons. The regression lines in both graphs appear to be significantly different for tight and wide branching.

### Compare manual vs. 3D model measurements

```{r}

```


### Visualize invertebrate data

```{r, message=FALSE} 
# Create two graphs showing invert abundance and coral volume for both tight and wide corals
invert_vol <- ggplot(cafi_coral2, aes(x=volume_pg, y=num_cafi, shape= branch, color= branch, linetype= branch))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Invert. abundance", x= expression(paste("Volume m"^{3})), title = "Invert. abundance and coral volume")

invert_vol
```

```{r, include=FALSE} 
#fit linear model for tight branching coral volume and invertebrate abundance
fit_invert_vol_tight <- lm(num_cafi~volume_pg, data=cafi_coral2, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_invert_vol_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE}
#fit linear model for wide branching coral volume and invertebrate abundance
fit_invert_vol_wide <- lm(num_cafi~volume_pg, data=cafi_coral2, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_invert_vol_wide)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  volume_pg = "Volume"
)
tab_model(fit_invert_vol_tight, fit_invert_vol_wide,
          pred.labels = pl,
          dv.labels = c("Invert abundance in tight coral", "Invert abundance in wide coral"))
```

##### Looking at the table for invertebrate abundance in both wide and tight corals, we find that none of the slopes are significantly different from zero. Based on the overlap of the confidence intervals, we cannot conclude that the slopes are signficantly different from eachother. Volume is not significantly correlated for tight or wide coral. 

```{r, message=FALSE} 
# Show invert abundance and coral SA for both tight and wide corals
invert_SA <- ggplot(cafi_coral2, aes(x=surface_area, y=num_cafi, shape= branch, color= branch, linetype= branch))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Invert. abundance", x= expression(paste("Surface area m"^{2})), title = "Invert. abundance and coral SA")

invert_SA
```

```{r, include=FALSE} 
#fit linear model for tight branching coral SA and invertebrate abundance
fit_invert_SA_tight <- lm(num_cafi~surface_area, data=cafi_coral2, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_invert_SA_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE}
#fit linear model for wide branching coral SA and invertebrate abundance
fit_invert_SA_wide <- lm(num_cafi~surface_area, data=cafi_coral2, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_invert_SA_wide)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  surface_area = "SA"
)
tab_model(fit_invert_SA_tight, fit_invert_SA_wide,
          pred.labels = pl,
          dv.labels = c("Invert abundance in tight coral", "Invert abundance in wide coral"))
```

##### Looking at the table relating invertebrate abundance to surface area, we can conclude that the slope for surface area and wide corals is significantly different from zero. There is a weak correlation for surface area and invertebrate abundance for wide and tight corals. 

```{r, message=FALSE}
# Scatter plot of average branch width and invertebrate abundance for both wide and tight corals
invert_branch <- ggplot(cafi_coral2, aes(x=avg_w_cm, y=num_cafi, shape=branch, color=branch, linetype=branch))+ #color=as.factor(branch)))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+ 
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Invert. abundance", x= "Avg. branch distance mm", title = "Invert. abundance and branch distance")

invert_branch
```

```{r, include=FALSE} 
#fit linear model for tight branching coral avg. branch distance and invertebrate abundance
fit_invert_branch_tight <- lm(num_cafi~avg_w_cm, data=cafi_coral2, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_invert_branch_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE}
#fit linear model for wide branching coral avg. branch distance and invertebrate abundance
fit_invert_branch_wide <- lm(num_cafi~avg_w_cm, data=cafi_coral2, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_invert_branch_wide)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  avg_w_cm = "Avg. branch width"
)
tab_model(fit_invert_branch_tight, fit_invert_branch_wide,
          pred.labels = pl,
          dv.labels = c("Invert abundance in tight coral", "Invert abundance in wide coral"))
```

##### Based on our linear regression and table for avg. branch width and invertebrate abundance, we cannot conclude that any of the slopes are significantly different from zero or different from each other. There is no correlation between avg. branch distance and invertebrate abundance.

##### For this step we use the *pairs.panels()* function to check for correlation between variables and check for normality. 
```{r, error==FALSE}
cafi_coral3 <- cafi_coral2 %>% select(num_cafi,volume_pg, avg_w_cm, surface_area) 
pairs.panels(cafi_coral3, lm = TRUE, cor = T) #surface area is highly correlated with branch distance and coral volume
cafi_coral3 <- cafi_coral3 %>% subset(select = -surface_area) #drop from dataset
```

```{r, include=TRUE}
#data is skewed right, log transform
logcafi_coral3 <- log(cafi_coral3)

logfit_1var <- lm(num_cafi ~ volume_pg, 
               data = logcafi_coral3)

logfit_2var <- lm(num_cafi ~ volume_pg + avg_w_cm,
               data = logcafi_coral3)

#Look at diagnostic plots for each model
par(mfrow=c(2,2))
plot(logfit_1var) #data does NOT appear to be normal
par(mfrow=c(1,1))

#Look at diagnostic plots for each model
par(mfrow=c(2,2))
plot(logfit_2var) #data appears somewhat normal
par(mfrow=c(1,1))
```

```{r}
pairs.panels(logcafi_coral3, lm = TRUE, cor = T)
```

```{r, include=FALSE}
#calculate AIC of each model
result <- AIC(logfit_1var,logfit_2var) #create a dataframe whose rownames are the models, with columns for the df and AIC of each model

#add other metrics to table
models <- list(logfit_1var,logfit_2var) #keep models in the same order here as they were when you created your results table
result$BIC <- sapply(models, BIC) #add a column for BIC to the results

model_summary <- lapply(models, summary) 

#use a for loop to easily extract the R^2 and adj R^2 value for each model from its summary, and store them in new columns in the results table

for(i in 1:length(models)){ #this creates a variable i that starts with the value i=1
  result$rsq[i] <- model_summary[[i]]$r.squared #we assign the rsq value from model i to the i'th row of the column 'rsq' in the table 'results'
  result$adj_rsq[i] <- model_summary[[i]]$adj.r.squared #same for adjusted rsq
} #now we go back to the beginning of the for-loop, add 1 to the value of i, and do everything again


kable(result, digits = 2, align = "c")
```

```{r, include=FALSE}
splitter <- sample(1:nrow(logcafi_coral3), 15, replace = F) #pick 15 random rows to reserve as test data
coral_train <- logcafi_coral3[-splitter,] #leave those rows out of the training data
coral_test <- logcafi_coral3[splitter,] #use them to create a set of test data

logfit_2var_split <- lm(num_cafi ~ volume_pg + avg_w_cm,
               data = logcafi_coral3) #fit the final model, using JUST your training set as the data argument

prediction <- predict(logfit_2var_split,coral_test) #use the fitted model to predict values for your test data

plot(coral_test$num_cafi, pch=1) #plot the actual test data values
points(prediction, pch=20, col = "red") #plot the model predictions for those points
```


```{r, warning=FALSE}
# Boxplot of invert abundance
invert_box <- ggplot(cafi_coral2, aes(x= branch, y=num_cafi, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter()+ # adds data on top of box plot
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(cafi_coral2$num_cafi + 3)), breaks = c(0, 10, 20, 30, 40, 50))+
  theme_classic()+
  labs(y="Invert. abundance", x="Coral branching", title = "Invert. abundance boxplot")

invert_box

# Boxplot of invertebrate species richness per coral for both tight and wide branching
richness_box <- ggplot(cafi_coral2, aes(x= branch, y=cafi_richness, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter()+
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(cafi_coral2$cafi_richness + 5)), breaks = c(0, 5, 10, 15, 20))+
  theme_classic()+
  labs(y="Invert. sp. richness", x="Coral branching", title = "Invert. sp. richness boxplot")

richness_box

# Boxplot of invertebrate diversity for both tight and wide branching coral
diversity_box <- ggplot(cafi_coral2, aes(x= branch, y=sw, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter()+
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(cafi_coral2$sw + 0.5)), breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3))+
  theme_classic()+
  labs(y="Invert. diversity", x="Coral branching", title = "Invert. diversity boxplot")

diversity_box
```


```{r, include=FALSE}
### Dataframe of Trapezid crab abundance and combine with photogrammetry coral data

# Calculate Trapezid abundance and richness for each coral
its_a_trap <- updated_cafi2 %>% filter(str_detect(code, "^TR"))
trap_summarized <- group_by(its_a_trap, coral_id) %>%
  summarise(num_cafi = n(), cafi_richness = length(unique(code)), cafi_present = paste(sort(unique(code)), collapse = ";"))
```

```{r, include=FALSE}
# Combine with photogrammetry data
trap_coral_dim <- merge(trap_summarized, coral_dim, by = "coral_id")
trap_coral_dim$num_trap <- trap_coral_dim$num_cafi %>% tidyr::replace_na(0) %>% rename("num_trap")
trap_coral_dim$trap_richness <- trap_coral_dim$cafi_richness %>% tidyr::replace_na(0) %>% rename("trap_richness")
trap_coral_dim <- trap_coral_dim %>% select(-cafi_richness, -num_cafi) %>% drop_na(volume_pg, surface_area)
```


### Visualize Trapezid data
```{r, message=FALSE}
# Scatter plot of volume and Trapezid abundance for both wide and tight corals
trap_vol <- ggplot(trap_coral_dim, aes(x=volume_pg, y=num_trap, shape=branch, color=branch, linetype=branch))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Trapezid abundance", x= expression(paste("Volume m"^{3})), title = "Trap. abundance and coral volume")

trap_vol
```

```{r, include=FALSE} 
#fit linear model for tight branching coral volume and trapezid abundance
fit_trap_vol_tight <- lm(num_trap~volume_pg, data=trap_coral_dim, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_trap_vol_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE} 
#fit linear model for wide branching coral volume and trapezid abundance
fit_trap_vol_wide <- lm(num_trap~volume_pg, data=trap_coral_dim, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_trap_vol_wide)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  volume_pg = "Volume"
)
tab_model(fit_trap_vol_tight, fit_trap_vol_wide,
          pred.labels = pl,
          dv.labels = c("Trap. abundance in tight coral", "Trap abundance in wide coral"))
```

##### Both linear regression lines are significantly different from zero for volume and Trapezid abundance. The lines are also signficantly different from eachother with moderate correlation between coral volume and abundance. 

```{r, message=FALSE} 
# Show invert abundance and coral SA for both tight and wide corals
trap_SA <- ggplot(trap_coral_dim, aes(x=surface_area, y=num_trap, shape= branch, color= branch, linetype= branch))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Trap. abundance", x= expression(paste("Surface area m"^{2})), title = "Trap. abundance and coral SA")

trap_SA
```
```{r, include=FALSE} 
#fit linear model for tight branching coral SA and invertebrate abundance
fit_trap_SA_tight <- lm(num_trap~surface_area, data=trap_coral_dim, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_trap_SA_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE}
#fit linear model for wide branching coral SA and invertebrate abundance
fit_trap_SA_wide <- lm(num_trap~surface_area, data=trap_coral_dim, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_trap_SA_wide)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  surface_area = "SA"
)
tab_model(fit_trap_SA_tight, fit_trap_SA_wide,
          pred.labels = pl,
          dv.labels = c("Invert abundance in tight coral", "Invert abundance in wide coral"))
```

##### Both linear regression slopes are significantly different from zero for both tight and wide corals. Confidence intervals overlap and further tests will have to be conducted to confirm if the slopes are signficantly different from eachother. There is a stronger correlation for SA and tight corals than for wide corals. 

```{r, message=FALSE}
# Scatter plot of average branch width and Trapezid abundance for wide and tight corals
trap_branch <- ggplot(trap_coral_dim, aes(x=avg_w_cm, y=num_trap, shape=branch, color=branch, linetype=branch))+
  geom_point()+
  geom_smooth(method = lm, se=TRUE, aes(fill=branch))+
  theme_classic()+
  #facet_wrap(~branch)+
  labs(y="Trapezid abundance", x="Avg. branch distance mm", title = "Trap. abundance and branch distance")

trap_branch
```

```{r, include=FALSE} 
#fit linear model for tight branching coral avg. branch distance and trapezid abundance
fit_trap_branch_tight <- lm(num_trap~avg_w_cm, data=trap_coral_dim, subset=(branch=="tight"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_trap_branch_tight)
par(mfrow=c(1,1))
```

```{r, include=FALSE} 
#fit linear model for wide branching coral avg. branch distance and trapezid abundance
fit_trap_branch_wide <- lm(num_trap~avg_w_cm, data=trap_coral_dim, subset=(branch=="wide"))
```

```{r, include=FALSE}
par(mfrow=c(2,2))
plot(fit_trap_branch_wide)
par(mfrow=c(1,1))
```

```{r}
#HTML table from regression results
pl <- c(
  '(Intercept)' = "Intercept",
  avg_w_cm = "Avg. branch distance"
)
tab_model(fit_trap_branch_tight, fit_trap_branch_wide ,
          pred.labels = pl,
          dv.labels = c("Trap. abundance in tight coral", "Trap abundance in wide coral"))
```

##### There is no significant difference from zero for the linear regression line slopes and they are not signficantly different from eachother. There is extremely weak correlation for avg. branch distance and Trapezid abundance for both tight and wide corals. 

```{r}
#Boxplot of Trapezid crab abundance
trap_box <- ggplot(trap_coral_dim, aes(x= branch, y=num_trap, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter(height = 0)+ # adds data on top of box plot
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(trap_coral_dim$num_trap + 5)), breaks = c(0, 5, 10, 15, 20))+
  theme_classic()+
  labs(y="Trapezid abundance", x="Coral branching", title = "Trap. abundance boxplot")

trap_box

#Boxplot of Trapezid species richness per coral for both tight and wide branching
trap_richness_box <- ggplot(trap_coral_dim, aes(x= branch, y=trap_richness, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter(height = 0)+ # adds data on top of box plot
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(trap_coral_dim$trap_richness + 1)), breaks = c(0, 1, 2, 3, 4, 5))+
  theme_classic()+
  labs(y="Trapezid species richness", x="Coral branching", title = "Trap. sp. richness boxplot")

trap_richness_box
```

```{r, include=FALSE}
### Dataframe of Alpheus shrimp abundance and combine with photogrammetry coral data

alph <- updated_cafi2 %>% filter(str_detect(code, "^AL"))
alph_summarized <- group_by(alph, coral_id) %>%
  summarise(num_cafi = n(), cafi_richness = length(unique(code)), cafi_present = paste(sort(unique(code)), collapse = ";"))
```

```{r, include=FALSE}
alph_coral_dim <- merge(alph_summarized, coral_dim, by = "coral_id")
alph_coral_dim$num_alph <- alph_coral_dim$num_cafi %>% 
  tidyr::replace_na(0) %>% 
  rename("num_alph")
alph_coral_dim$alph_richness <- alph_coral_dim$cafi_richness %>% 
  tidyr::replace_na(0) %>% 
  rename("alph_richness")
alph_coral_dim <- alph_coral_dim %>% 
  select(-cafi_richness, -num_cafi) %>% 
  drop_na(volume_pg, surface_area)
```

### Visualize Alpheus shrimp data

##### There were not enough data points for Alepheus shrimp and we could not make predictions for shrimp abundance based on the high degree of scatter. Plots were included as a visual representation but will not be analyzed. We will wait for more species to be identified and further data cleaning.

```{r, message=FALSE}
# Scatter plot of volume and Alpheus shrimp abundance for both wide and tight corals
alph_volume <- ggplot(alph_coral_dim, aes(x=volume_pg, y=num_alph))+
  geom_point()+
  geom_smooth(method = lm, se=FALSE, color = "red")+
  theme_classic()+
  facet_wrap(~branch)+
  labs(y="Alpheus abundance", x= expression(paste("Volume m"^{3})), title = "Alpheus shrimp abundance and coral volume")

alph_volume
```

```{r, message=FALSE}
# Scatter plot of average branch width and Trapezid abundance for both wide and tight corals
alph_branch <- ggplot(alph_coral_dim, aes(x=avg_w_cm, y=num_alph))+
  geom_point()+
  geom_smooth(method = lm, color = "red", se=FALSE)+
  theme_classic()+
  facet_wrap(~branch)+
  labs(y="Alpheus abundance", x="Avg. branch distance mm", title = "Alpheus shrimp abundance and average branch width of coral")

alph_branch
```

```{r}
# Boxplot of Alpheus shrimp abundance
alph_box <- ggplot(alph_coral_dim, aes(x= branch, y=num_alph, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter(height = 0.25)+ # adds data on top of box plot
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(alph_coral_dim$num_alph + 1)), breaks = c(0, 1, 2, 3, 4))+
  theme_classic()+
  labs(y="Alpheus abundance", x="Coral branching", title = "Alpheus shrimp abundance for tight and wide branching coral")

alph_box

# Boxplot of Alpheus shrimp species richness per coral for both tight and wide branching
alph_richness_box <- ggplot(alph_coral_dim, aes(x= branch, y=alph_richness, color = branch))+
  geom_boxplot(width = 0.5)+
  geom_jitter(height = 0.25)+ # adds data on top of box plot
  scale_color_manual(values = c("firebrick1", "steelblue2"))+
  scale_y_continuous(limits= c(0, max(alph_coral_dim$alph_richness + 1)), breaks = c(0, 1, 2, 3))+
  theme_classic()+
  labs(y="Alpheus species richness", x="Coral branching", title = "Alpheus shrimp species richness for tight and wide branching coral")

alph_richness_box
```
